---
title: "EDA"
---

# I made statistics on the number of electric cars and hybrid electric cars purchased around the world from 2010 to 2022. The following is a specific analysis.

```{r}
df=read.csv("/Users/kenny/Desktop/dsan-5000-project-kennyzhao0428/data/01-modified-data/EVsalesHistori.csv")
# Trend Analysis: Yearly EV Sales and Stock
library(ggplot2)

# Total sales
ev_sales_trend <- ggplot(df[df$parameter == "EV sales", ], aes(x = year, y = value, group = powertrain, color = powertrain)) +
  geom_line() +
  labs(title = "EV Sales Trend Over Years", x = "Year", y = "Sales Volume")

# Total stock
ev_stock_trend <- ggplot(df[df$parameter == "EV stock", ], aes(x = year, y = value, group = powertrain, color = powertrain)) +
  geom_line() +
  labs(title = "EV Stock Trend Over Years", x = "Year", y = "Stock Volume")

# Print the plots
print(ev_sales_trend)
print(ev_stock_trend)
```

# The two charts above reflect the significant increase in the number of trams purchased since 2010, and also reflect the increasing acceptance of trams. Especially after 2019, the number of people buying trams has increased several times.

```{r}
# Proportion Analysis: Yearly EV Sales Share
ev_sales_share <- ggplot(df[df$parameter == "EV sales share", ], aes(x = year, y = value, fill = powertrain)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "EV Sales Share Over Years", x = "Year", y = "Sales Share (%)")

# Print the plot
print(ev_sales_share)
```


# The following is an analysis of the correlation of this set of data。

```{r}
# Correlation Analysis: Stock vs. Sales
# You first need to create a new data frame that only has the sales and stock data
sales_data <- df[df$parameter == "EV sales", c("year", "powertrain", "value")]
stock_data <- df[df$parameter == "EV stock", c("year", "powertrain", "value")]

# Rename columns for merging
colnames(sales_data)[3] <- "sales"
colnames(stock_data)[3] <- "stock"

# Merge the data frames
merged_data <- merge(sales_data, stock_data, by = c("year", "powertrain"))

# Calculate correlation
cor_matrix <- cor(merged_data[, c("sales", "stock")], use = "complete.obs")

# Display the correlation matrix
print(cor_matrix)
```

# We can see that this set of data is still very good. and we can also do the outlier as follow:

```{r}
Q1 <- quantile(df$value, 0.25)
Q3 <- quantile(df$value, 0.75)
IQR <- Q3 - Q1
# Define the bounds for outliers
lower_bound <- Q1 - 1.5 * IQR
upper_bound <- Q3 + 1.5 * IQR
# Identify the outliers
outliers <- df$value < lower_bound | df$value > upper_bound
# Filter the outlier data
outliers_data <- df[outliers, ]
# Visualize the outliers
boxplot(df$value, main="Boxplot for 'value' Column")
```



# For the data about：Men and women of different age groups consider the cost of purchasing electric vehicles, the proportion of electric vehicles purchased, and the concerns that influence their purchase of electric vehicles.

```{r}
library(dplyr)
library(ggplot2)
library(corrplot)

df = read.csv("/Users/kenny/Desktop/dsan-5000-project-kennyzhao0428/data/01-modified-data/clean26.csv")
str(df)

# delete “Total males” and “Total females”
df_filtered = df[!df$Sex.Age.group %in% c("Total males", "Total females"), ]

# mean, median and sd

df_stats = df_filtered %>%
  summarise(
    Mean_Cost = mean(Cost, na.rm = TRUE),
    Median_Cost = median(Cost, na.rm = TRUE),
    SD_Cost = sd(Cost, na.rm = TRUE)
  )

print(df_stats)

# visulize
ggplot(df_filtered, aes(x = Cost)) + 
  geom_histogram(binwidth = 5, fill = "blue", color = "black") +
  theme_minimal() +
  labs(title = "Cost Distribution", x = "Cost", y = "Frequency")

# boxplot
ggplot(df_filtered, aes(x = `Sex.Age.group`, y = Cost)) + 
  geom_boxplot() + 
  theme_minimal() +
  labs(title = "Cost by Age Group", x = "Age Group", y = "Cost")

# Correlation Analysis
grouped_data <- df_filtered %>%
  group_by(`Sex.Age.group`) %>%
  summarise(Mean_Cost = mean(Cost, na.rm = TRUE))

print(grouped_data)

# outlier
boxplot_stats <- boxplot(Cost ~ `Sex.Age.group`, data = df_filtered, outline = TRUE, main = "Cost by Age Group")
```

