
## Clean up the data about proportion of electric cars owned by people of different genders and ages

[can find the data here](https://www.cso.ie/en/releasesandpublications/ep/p-smt/sustainablemobilityandtransport2021/electricvehicles/)

```{r}
library(readxl)
library(tidyverse)

#Read in the Data
data = read_excel("/Users/kenny/Desktop/dsan-5000-project-kennyzhao0428/dsan-website/5000-website/82.xlsx", skip = 2)

#Remove Unwanted Rows
df = data %>%
  mutate(Gender = case_when(
    `...1` == "Male" ~ "Male",
    `...1` == "Female" ~ "Female",
    TRUE ~ NA_character_
  ))
df = df %>%
  fill(Gender)

#Remove unnecessary rows
df = df %>%
  filter(!(`...1` %in% c("Male", "Female", "All Persons", "Sex/Age")))

#Fix the accuracy of numbers
df = df %>%
  mutate(`Electric Vehicle (EV) Ownership` = round(as.numeric(`Electric Vehicle (EV) Ownership`), 1))

#Rename columns
df = df %>%
  rename(
    `Age Group` = `...1`,
    `EV ownership Yes` = `Electric Vehicle (EV) Ownership`,
    `EV ownership No` = `...3`,
    Total = `...4`
  )

#Delete useless lines
df = df %>%
  slice(-c(1, 9))

print(df)
```



```{r}
library(readxl)
library(tidyverse)

#Read in the Data
data = read_excel("/Users/kenny/Desktop/dsan-5000-project-kennyzhao0428/dsan-website/5000-website/26.xlsx", skip = 2)

library(dplyr)

#Delete useless lines
df = data %>% filter(!apply(is.na(.), 1, all))

#Fix the accuracy of numbers
df = df %>%
  mutate(`...3` = round(as.numeric(`...3`), 1))%>%
  mutate(`...4` = round(as.numeric(`...4`), 1))%>%
  mutate(`...5` = round(as.numeric(`...5`), 1))%>%
  mutate(`...6` = round(as.numeric(`...6`), 1))%>%
  mutate(`...7` = round(as.numeric(`...7`), 1))%>%
  mutate(`...8` = round(as.numeric(`...8`), 1))%>%
  mutate(`...9` = round(as.numeric(`...9`), 1))%>%
  mutate(`...10` = round(as.numeric(`...10`), 1))%>%
  mutate(`...11` = round(as.numeric(`...11`), 1))%>%
  mutate(`...12` = round(as.numeric(`...12`), 1))

#Delete useless lines
df = df[-(17:20), ]

#rename
names(df)[1:ncol(df)] = c("Sex/Age group","Cost", "Inadequete public charging facilities", "No/limited access to home charging", "Limited range available", "Don't drive/ own a car", "Resale value of EV", "Reliability/ safety concerns", "No need/ don't know enough", "Not changing for a while", "Not available at time of last purchase", "Other", "Gender")

#Fix the accuracy of numbers
df = df %>%
  mutate(`Cost` = round(as.numeric(`Cost`), 1))

#Reshaping gender data
df = df %>%
  mutate(Gender = case_when(
    `Sex/Age group` == "Male" ~ "Male",
    `Sex/Age group` == "Female" ~ "Female",
    TRUE ~ NA_character_
  ))
df = df %>%
  fill(Gender)

#Delete useless lines
df = df %>%
  slice(-c(1, 3, 10))

print(df)
```

```{r}
df=read.csv("/Users/kenny/Desktop/dsan-5000-project-kennyzhao0428/data/01-modified-data/EVsalesHistori.csv")

```

After cleaning and examining the dataset titled 'EVsalesHistori.csv', we can observe several key aspects that offer insights into the historical sales and market penetration of Electric Vehicles (EVs) in various regions, specifically focusing on Australia as an example.

The dataset is organized into several columns, each representing a different dimension of the EV market. The 'region' column indicates the geographical area of the data, with the initial rows showing data for Australia. The 'category' column is labeled as 'Historical', suggesting that the data reflects past trends and figures.

One of the main parameters measured is the 'EV stock', which refers to the total number of EVs in the region. In 2011, for example, Australia had an EV stock of 49 vehicles. This is complemented by the 'EV stock share', which represents the proportion of EVs in the overall vehicle market, a crucial indicator of market penetration. In 2011, this share was a mere 0.00046 percent in Australia, highlighting the nascent stage of EV adoption at that time.

Another critical parameter is the 'EV sales share', showing the percentage of EVs among all vehicle sales. In Australia in 2011, this share was 0.00650 percent. The 'EV sales' parameter directly shows the number of EVs sold, with 49 vehicles sold in 2011. The data also differentiates between types of EVs, such as Battery Electric Vehicles (BEV) and other electric powertrains.

Overall, the dataset provides a comprehensive view of the EV market's historical development. It reveals the gradual increase in the adoption of EVs over the years, demonstrating the growing consumer interest and the evolving automotive market dynamics. The data can be instrumental for policymakers, market analysts, and industry stakeholders in understanding past trends to forecast future developments in the EV industry. ​